---
title: "DATA 605 - Assignment 5 - Least Squares"
author: "Christopher Martin"
date: "September 25, 2016"
output:
  html_document:
    fig_caption: yes
    force_captions: yes
    highlight: pygments
    number_sections: yes
    theme: united
    toc: no
  pdf_document:
    fig_caption: yes
    highlight: pygments
    latex_engine: xelatex
    number_sections: yes
    toc: no
title2: Least Squares
---

# Problem Set 1

n this problem set we'll work out some properties of the least squares solution that we reviewed in the weekly readings.  Consider the unsolvable system **Ax** = **b** as given below

$$
\left(\begin{matrix}
1 & 0\\
1 & 1\\
1 & 3\\
1 & 4
\end{matrix}\right)
\left(\begin{matrix}
x1 \\
x2
\end{matrix}\right)
=
\left(\begin{matrix}
0\\
8\\
8\\
20
\end{matrix}\right)
$$

Write code in R to compute $X = (A)(A^T)$ and $Y = (A^T)(A)$. Then, compute the eigenvalues and eigenvectors of **X** and **Y** using the built-in commands in R.

```{r}
A <- matrix(c(1,0,1,1,1,3,1,4),nrow=4,byrow=TRUE)
A

#X = A * A^T
X <- A %*% t(A)
X

#Y = A^T * A
Y <- t(A) %*% A
Y

#x = eigenvalue of X
x <- eigen(X)
x <- x$val
round(x,4)

#y = eigenvalue of Y 
y <- eigen(Y)
y <- y$val
y
```

a) Write R markdown script to compute **(A^T)(A)** and **(A^T)(b)**

```{r}
#X = A^T * A
X

b = matrix(c(0,8,8,20), nrow=4)
b

#Z = A^T * b
Z <- t(A) %*% b
Z
```

b) Solve for x(carrot) in R using the above two computed matrices

```{r}
#Ax = b
A
b

#compute the SVD of A
#A = USV^T
asvd <- svd(A)
asvd

#this step was not required, I just wanted to see the result
#check that A = USV^T
round(asvd$u %*% diag(asvd$d) %*% t(asvd$v),0)
A

#solving for x(carrot)
#x = V * S * (U^T * b)
S <- diag(1/asvd$d)
S

x <- asvd$v %*% S %*% t(asvd$u) %*% b
x

check <- A %*% x
check
b
```

Not surprisingly (since we were told this in the prompts), this matrix isn't solveable but in turn there is a residual present in which R = ||Ax - b||:

c) what is the squared error of this solution?

```{r}
#e = b - Ax
e <- b - A %*% x
e

#check the solution
#Ax + e = b
A%*%x + e
b
```

d) Instead of **b** = [0;8;8;20], start with **p** = [1;5;13;17] and find the exact solution (i.e. show that this system is solvable as all equations are consistent with each other. This should result in an error vector **e** = **0**).

```{r}
p = matrix(c(1,5,13,17), nrow=4, byrow=TRUE)
p

x <- asvd$v %*% S %*% t(asvd$u) %*% p
x

check <- A %*% x
check
p
```

e) Show that the error **e** = **b** - **p** = [-1;3;-5;3].

```{r}
#e = b - p = [-1;3;-5;3]
e
b - p
```


f) Show that the error **e** is orthogonal to **p** and to each of the columns of **A**

```{r}
#check that e is orthogonal to p
#dot product should equal 0
t(e) %*% p

#check that e is orthogonal to each column of A
#dot product should equal 0
t(e) %*% A[,1]
t(e) %*% A[,2]
```


# Problem Set 2

Consider the modified auto-mpg data (obtained from the UC Irvine Machine Learning dataset). This dataset contains 5 columns 1) displacement, 2) horsepower, 3) weight, 4) acceleration, 5) mpg. We are going to model mpg as a function of the other four variables.

Write an R markdown script that takes in the auto-mpg data, extracts an A matrix from the first 4 columns and b vector from the fifth (mpg) column. Using the least squares approach, your code should compute the best fitting solution. That is, find the best fitting equation that expresses mpg in terms of the other 4 variables. Finally, calculate the fitting error between the predicted mpg of your model and the actual mpg. Your script should be able to load in the 5 column data set, extract A and b, and perform the rest of the calculations. Please have adequate comments in your code to make it easy to follow your work.

```{r}
#import auto-mpg data
autompgdata <- read.table('C:/Users/cgmartin/Documents/CUNY MSDA/DATA605/auto-mpg.data')
head(autompgdata)

#create matrix A -- first four column (displacement, horsepower, weight, acceleration)
matrixA <- as.matrix(autompgdata[,1:4])
head(matrixA)

#create matrix b -- fifth column (mpg)
matrixb <- matrix(autompgdata[,5])

#created function to fit the equation
myfit <- function(variablematrix, resultmatrix){
  asvd <- svd(variablematrix)
  S <- diag(1/asvd$d)
  x <- asvd$v %*% S %*% t(asvd$u) %*% resultmatrix
  x
}

matrixx <- myfit(matrixA,matrixb)

#fitting error
#e = b - Ax
matrixe <- matrixb - matrixA %*% matrixx
head(matrixe)

#check the solution
#Ax + e = b
head(matrixA %*% matrixx + matrixe)
head(matrixb)
```


Adding test cases to demonstrate that your code is working will be very helpful.

```{r}
#reproducability, random number used (123)
set.seed(123)

#create new test matrix A
matrixA2 <- matrix(c(rnorm(140, mean = mean(matrixA[,1]), sd = sd(matrixA[,1])),rnorm(140, mean = mean(matrixA[,2]), sd = sd(matrixA[,2])),rnorm(140, mean = mean(matrixA[,3]), sd = sd(matrixA[,3])),rnorm(140, mean = mean(matrixA[,4]), sd = sd(matrixA[,4]))),ncol=4)

#create new test matrix b
matrixb2 <- matrix(c(rnorm(140, mean = mean(matrixb))),ncol=1)

matrixx2 <- myfit(matrixA,matrixb)

#fitting error
#e = b - Ax
matrixe2 <- matrixb2 - matrixA2 %*% matrixx2
head(matrixe2)

#check the solution
#Ax + e = b
head(matrixA2 %*% matrixx2 + matrixe2)
head(matrixb2)
```

